### 2013-01-06 edupar-13/conclusion.tex
Incorporate nwf's feedback

\section{Future Work and Conclusion}
\label{sec:future}

we did stuff

\section*{Acknowledgments}

Many people have contributed to the development of
15-410 over the past decade.
Stephen Muckle, currently at Qualcomm, led
the switch to x86 projects running in Simics.
Nathaniel Wesley Filardo,
currently a Ph.D.\ candidate at Johns Hopkins,
increased the sanity of the build infrastructure,
designed the \x{swexn} system call, and helped critique this paper.
The current reference kernel is named ``pathos''
and was written by Michael J. Sullivan,
currently a CMU CS Ph.D.\ candidate,
and Elly Fong-Jones, currently at Google.
Over the past decade the course has benefitted
greatly from free educational Simics licenses,
donated initially by Virtutech and recently by
Wind River.