### 2013-01-09 edupar-13/intro.tex
intro.tex: Clarification pass.

\section{Introduction}

Managing parallel computation, whether on thousand-node clusters or on one single-core machine, is often taught by resorting to powerful abstractions that simplify the hardest problems.
Programming patterns such as SPMD, divide-and-conquer, and fork/join parallelism enable developers to reason about parallel algorithms at a high, abstract level.
The user-level libraries and distributed systems that provide these abstractions, in turn, rely on operating system kernels to simplify the hardware and communication mechanisms on which they are built. % garth's

With the advent of massively multicore hardware,
kernels themselves have increasing hardware concurrency and complexity. % garth's
% there is also a need for reasoning about parallelism at lower levels: those of operating system kernels and concurrent systems applications, which are built on (or even below) the abstractions of threads and processes. % ben's old text
Low-level systems programmers must reason about concurrent code without the convenience of high-level simplifications, and hence often encounter notoriously difficult concurrency errors.

In this paper we describe our experience teaching concurrent programming at the operating systems level to advanced undergraduate students at Carnegie Mellon University (CMU).
In 15-410, the Operating Systems Design and Implementation class at CMU, students complete an aggressively low-level kernel project, which runs on real hardware, in which they implement threads and processes.
Presented with a kernel specification, called Pebbles~\cite{kspec}, students build a threading and synchronization library that runs in user space on top of a Pebbles-compliant kernel in a two-week project, and subsequently, in a six-week project, write their own multi-threaded kernel implementation.
During these projects students face demanding synchronization problems, including how to implement locking primitives and how to coordinate the possibly-simultaneous exiting of parent and child processes.
Finally, in another two-week project, students implement a small extension to their Pebbles kernel. This project changes every semester, and in the past has included extending their kernel serve as a hypervisor, or to support symmetric multiprocessing.

As we ask our students to write more and more difficult concurrent code in one semester, we wish to ease their burden of debugging concurrency errors, which often consumes a significant fraction of their time.
However, we also want to avoid compromising the complexity, hardware fidelity, and learning experience of the project set.
Hence, we have built Landslide~\cite{landslide}, a tool that uses systematic exploration~\cite{verisoft} to force kernel threads to take execution paths that expose concurrency bugs.
We present our experience sharing a prototype of Landslide with volunteering students who used it to find bugs in their own code, and discuss the possibility for integrating it as a main component of the 15-410 curriculum.

% should we do a "The contributions of this paper are as follows:"?

The rest of this paper is organized as follows.
Section~\ref{sec:curriculum} gives an overview of 15-410's five programming projects.
Section~\ref{sec:pebbles} discusses the Pebbles kernel project and how it is graded.
Section~\ref{sec:landslide} describes Landslide's testing mechanisms and how we worked with students to evaluate Landslide.
Section~\ref{sec:future} concludes, discussing strategies for making Landslide more accessible to struggling students, and how Landslide might be applied to related operating systems projects at other universities.