### 2015-11-03 pldi-16/paper.tex

%
% LaTeX template for prepartion of submissions to PLDI'15
%
% Requires temporary version of sigplanconf style file provided on
% PLDI'15 web site.
%
\documentclass[pldi]{sigplanconf-pldi15}

%
% the following standard packages may be helpful, but are not required
%
\usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{amsthm}
%\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI
%\doublespacing

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}

\begin{document}

%
% any author declaration will be ignored  when using 'plid' option (for double blind review)
%

% TODO: Don't forget to run e.g. s/\\quicksand /\\quicksand~/
\newcommand\landslide{\textsc{Landslide}}
\newcommand\quicksand{\textsc{Quicksand}}
\newcommand\simics{\textsc{Simics}}
\newcommand{\sect}[1]{\S #1}
\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{olivegreen}{RGB}{0,127,0}
\definecolor{brickred}{RGB}{192,0,0}

\newcommand\numthrlibs{79}
\newcommand\numpintoses{78} % 'di' is basecode. durr
\newcommand\numstudence{157} % total pintoses plus p2s

\title{Stateless Model Checking with Data-Race Preemption Points}
\authorinfo{Ben Blum}{Carnegie Mellon University}{bblum@cs.cmu.edu}
\authorinfo{Garth Gibson}{Carnegie Mellon University}{garth@cs.cmu.edu}

\maketitle
\begin{abstract}
Stateless model checking is a promising technique for testing concurrent programs,
but suffers from exponential explosion of the state space when the test input parameters are too large.
Several partial-order reduction techniques exist for mitigating this explosion,
but even after pruning equivalent interleavings, the state space size for a fixed set of preemption points is unpredictable and often intractable.
%
Data race detection, another concurrency testing approach, focuses on identifying suspicious memory access pairs during a single test execution.
It avoids concerns of intractable state space size, but is prone to false positives.

We present an {\em Iterative Deepening} framework for stateless model checking,
which manages the exploration of many state spaces using different subsets of preemption points.
It uses state space estimation to prioritize jobs most likely to complete in a fixed CPU budget,
and it incorporates a data-race analysis to dynamically add new preemption points to classify each data race as buggy or benign.
%
Our evaluation shows that combining data-race detection with stateless model checking can find new bugs neither technique would find on their own.
%Our evaluation shows this technique is
%more effective than single-state-space model checking
%%both at finding more bugs and at completing more state spaces when no bug exists.
%both at finding the same bugs faster and at finding new bugs entirely.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{intro}
\input{di}
\input{evaluation}
\input{related}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}

%We are great. Accept our paper.

We have presented \quicksand, an implementation of Iterative Deepening, a new technique for automating the choice of PPs during stateless model checking.
% , and \quicksand, a tool which implements this technique tailored to the \landslide~model checker.
\quicksand~incorporates \landslide's data-race analysis to create new PPs tailored specifically to the program under test,
and manages multiple \landslide~instances to test many different PP subsets in a given CPU budget, even when the full state space of all PPs would be computationally intractable.

Our evaluation shows that \quicksand~achieves better bug-finding results than either single-pass data-race detection or single-state-space model checking (SSS-MC) alone.
We dramatically reduce false-positive data race reports,
%by verifying the absence of bugs in their associated state spaces
and find bugs caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
% TODO: Not true; Replace instead with a claim about partial verification.
Compared to SSS-MC, we find bugs faster in smaller subset state spaces,
%when the maximal state space is too large to explore
and find new bugs with data-race PPs that the maximal state space would not expose at all.

%Although \landslide~itself is not publicly available due to its dependency on the non-free simulator \simics, \quicksand~is open-source and its interface can be adapted to fit any similar tool.
\quicksand~is open-source and its interface can be adapted to fit any MC tool similar to \landslide.
We have also posted our evaluation's full data set and logs.
They can be found here:

% TODO: Make data set, at least, public.
TODO: Data set at least can be public.

{\em github links removed for double-blind review}
%\url{https://github.com/bblum/landslide} % POST-SUBMISSION TODO: put QS in its own repository
%
%\url{https://github.com/bblum/landslide} % POST-SUBMISSION TODO: put final logs somewhere
% POST-SUBMISSION TODO: Don't forget to anonymize the names of all your grupos.

\section{Acknowledgements}

%Many thanks to David A. Eckhardt, Vaishaal Shankar, and Haryadi Gunawi for generously providing student implementations from CMU's, Berkeley's, and U. of Chicago's OS classes respectively.
Many thanks to Vaishaal Shankar, Haryadi Gunawi, and David A. Eckhardt for generously providing student implementations from Berkeley's, U. of Chicago's, and CMU's OS classes respectively.
Thanks to Wind River for the use of their simulator \simics.
Thanks to
% TODO: Post-double-blind-review
% Ji\v{r}\'{i} \v{S}im\v{s}a, John Wilkes, and
the anonymous PLDI reviewers for their helpful comments.
This work was supported in part by % TODO: funding sources
the U.S. Army Research Office under grant number W911NF0910273.

\bibliographystyle{abbrvnat}
\bibliography{citations}{}

\end{document}