### 2015-11-10 pldi-16/proof/convergence.tex

\section{Proof of Iterative Deepening convergence}

We seek to prove that, given enough time, Iterative Deepening offers the same coverage of thread interleavings that could be achieved by preempting at every single instruction.
We'll call the latter the {\em na\"{i}ve state space}, and call the condition of testing all its interleavings {\em convergence}.
(It could also be called soundness, from the perspective of viewing Iterative Deepening as a search re-ordering heuristic that doesn't miss any interleavings.)
Hence, our convergence statement is as follows:

\begin{theorem}[Total verification]
	If Iterative Deepening fully saturates its set of data-race PPs and completes all associated state spaces,
	it serves as a verification of all possible thread schedules of the given test program.
	\label{thm:totalverif}
\end{theorem}

The contrapositive statement offers more structure for an easier proof\footnote{
For the reader who likes to avoid non-constructive proofs, % \cite{vargomax}, % TODO: uncomment for camready/TR
note that we are using the constructive contrapositive direction:
Theorem~\ref{thm:convergence} is $A \rightarrow B$ and Theorem~\ref{thm:totalverif} is $\neg B \rightarrow \neg A$.}:

%DPOR + Iterative Deepening will eventually test every nonequivalent thread interleaving that would be exposed by preempting at every single instruction
\begin{theorem}[Convergence of Iterative Deepening]
	If a bug is exposed by an interleaving in the na\"{i}ve state space, Iterative Deepening will eventually test an equivalent interleaving which exposes the same bug.
	\label{thm:convergence}
\end{theorem}

%First we will prove a slightly weaker statement, limiting the na\"{i}ve approach's PPs to data races,
%then strengthen it by showing that the all interleavings in the na\"{i}ve state space
%are equivalent to some interleaving made of data-race PPs (and sync API PPs) alone.

%%%%

\newcommand\ppnext[1]{\ensuremath{\mathsf{next}(#1)}}
\newcommand\ppinstr[1]{\ensuremath{\mathsf{instr}(#1)}}
\newcommand\ppothers[1]{\ensuremath{\mathsf{others}(#1)}}
\newcommand\pppfx[1]{\ensuremath{\mathsf{pfx}(#1)}}
\newcommand\conflicts[2]{\ensuremath{\mathsf{conflicts}(#1,#2)}}
\newcommand\pai{\ensuremath{p_{\alpha{}i}}}
Let
\[
	\mathcal{I} = \{(t_{\alpha{}1}, p_{\alpha{}1}), (t_{\beta{}1}, p_{\beta{}1}), ... (t_{\alpha{}n}, p_{\alpha{}n}), ...\}
\]
be an interleaving trace, where the element $(t_{\alpha{}i}, \pai)$ represents the $i$th transition $t$ of thread $\alpha$, and the associated preemption point $p$.
We will use the following notation:
\begin{itemize}
	\item $\ppnext{p_{\alpha{}i}}$ indicates the next transition by the same thread, $t_{\alpha{}(i+1)}$,
	\item $\ppinstr{p_{\alpha{}i}}$ indicates the first instruction executed during $t_{\alpha{}(i+1)}$. and
	\item $\ppothers{p_{\alpha{}i}}$ indicates the set of all transitions by other threads between $t_{\alpha{}i}$ and $t_{\alpha{}(i+1)}$.
	\item $\conflicts{t_\alpha,t_\beta}$ indicates the set of memory access pairs between $t_\alpha$ and $t_\beta$ such that among each pair, one access belongs to each thread $\alpha$ and $\beta$, the address is the same, and at least one is a write; i.e., precisely the shared memory conflict relation required by DPOR to determine dependent interleavings.
\end{itemize}

\subsection{Equivalence of irrelevant PPs}

Iterative Deepening will not, of course, preempt on exactly the same instructions that an arbitrary naive interleaving would,
as it is only capable of preempting on data races and sync API boundaries.
Thus, our first task is to show that all naive interleavings have an equivalent interleaving with only data-race and sync API PPs.

% TODO: come up with a better name
\begin{definition}[Relevant preemption point]
	We say a PP $\pai$ is {\em relevant} if either $\ppnext{\pai}$ is a sync API boundary (including a voluntary yield),
	or if
	\[
		\conflicts{\ppothers{\pai}}{\ppinstr{\pai}} \ne \emptyset
	\]
	i.e., the instruction by thread $\alpha$ immediately after $\pai$ has a memory conflict with some other thread interleaved during $\pai$.
\end{definition}
\begin{definition}[Fully-relevant interleaving]
	An interleaving comprised only of relevant PPs.
\end{definition}

\begin{lemma}
	For any interleaving in the na\"{i}ve state space, there exists an equivalent interleaving which uses only relevant PPs.
	%data-race PPs and sync API PPs.
	\label{lem:equivalent}
\end{lemma}

\begin{proof}
	Let $\pai$ be the first irrelevant PP of a naive interleaving $\mathcal{I}$.
	We ask, did some thread among $\ppothers{\pai}$ conflict with any instruction from $\ppnext{\pai}$, even though there was no conflict with $\ppinstr{\pai}$?
	\begin{itemize}
			% rho for relevant.
		\item If $\conflicts{\ppothers{\pai}}{\ppnext{\pai}} \ne \emptyset$, or if $t_{\alpha{}i}$ contains a sync API boundary instruction that's not already a PP,
			then let $\rho$ be either the first instruction by $\ppnext{\pai}$ among the conflicts or the first sync API instruction, whichever comes first.
			Let $\pppfx{\rho}$ denote the instruction sequence between $\ppinstr{\pai}$ and $\rho$, including the former but not the latter.
			Now, we output the new interleaving:

			\begin{tabular}{lll}
				\\
				$\mathcal{I}' = \{...,$&$(t_{\alpha{}i} \cup \pppfx{\rho}, \pai'),$& \\
									   &$\ppothers{\pai},$& \\
					       &$(t_{\alpha{}(i+1)}) \setminus \pppfx{\rho}, p_{\alpha{}(i+1)}),$ & $ ...\}$ \\
				\\
			\end{tabular}

			In other words, we have simply reordered $\ppothers{\pai}$ to between $\pppfx{\rho}$ and $\rho$, removing the irrelevant $\pai$ and adding a new PP $\pai'$, which is relevant by the construction of $\rho$.
			We know it must be possible to reorder $\ppothers{\pai}$ to after $\alpha$'s execution because no synchronization is done during $\pppfx{\rho}$, hence the other threads' runnability cannot be affected.
			Likewise $\pppfx{\rho}$ is independent with $\ppothers{\pai}$, so by the soundness of DPOR \cite{dpor}, $\mathcal{I}' \equiv \mathcal{I}$.
		\item Otherwise, $\conflicts{\ppothers{\pai}}{\ppnext{\pai}} = \emptyset$, and no instructions among $\ppnext{\pai}$ are a sync API boundary.
			Then we output the new interleaving:
			\[
				\mathcal{I}' = \{..., (t_{\alpha{}i} \cup t_{\alpha(i+1)}, p_{\alpha(i+1)}), \ppothers{\pai},...\}
			\]
			In other words, we have reordered {\ppothers{\pai}} with the entire next transition by thread $\alpha$.
			We know this must be possible for the same reasons as above, and again, the transitions are independent, so $\mathcal{I}' \equiv \mathcal{I}$.
					       %&$(t_{\alpha{}(i+1)}) \setminus \pppfx{\rho}, p_{\alpha{}(i+1)}),$ & $ ...\}$ \\
	\end{itemize}
	This constitutes an algorithm for inductively converting (or removing) all irrelevant PPs to relevant ones.
\end{proof}

%%%%

\subsection{Saturation of data-race PPs}

Now we must show that, starting from a sync-PP-only state space,
Iterative Deepening will eventually interleave threads sufficiently to expose all data races which are used as PPs in any buggy fully-relevant interleaving.
The challenge of this proof is that some data-race candidates are nondeterministic to find;
i.e., they may be hidden in an obscure control flow path which requires a prior preemption on a different data race to expose,
as shown in Figure~\ref{fig:nondet-dr}.
Hence, the maximal state space of statically-available sync API PPs will not necessarily uncover all possible data-race PPs;
Iterative Deepening may need to iterate through some data-race state spaces before finding certain nondeterministic races.

\begin{figure}[t]
	\small
	\begin{tabular}{rl}
	& \multicolumn{1}{c}{~\texttt{int x = 0, y = 0;~~~~~~~~~~~~~~~~}} \\
	& \multicolumn{1}{c}{\texttt{bool t1\_x = false, t1\_y = false;}} \\
	& \multicolumn{1}{c}{\texttt{bool t2\_x = false, t2\_y = false;}} \\
\\
		& Thread 1 (Thread 2 similar, with {\tt t1} and {\tt t2} vars swapped) \\
	1 & \texttt{\hilight{brickred}{x = x + 1;}} \\
	2 & \texttt{t1\_x = true;} \\
	3 & \texttt{// "if x raced"} \\
	4 & \texttt{if (x == 1 \&\& t2\_x) \{} \\
	5 & \texttt{~~~~\hilight{brickred}{y = y + 1;}} \\
	6 & \texttt{~~~~t1\_y = true;} \\
	7 & \texttt{~~~~// "if y raced"} \\
	8 & \texttt{~~~~if (y == 1 \&\& t2\_y) \{} \\
	9 & \texttt{~~~~~~~~panic();} \\
	10 & \texttt{~~~~\}} \\
	11 & \texttt{\}} \\
	\end{tabular}
	\caption{Not all data races will immediately be uncovered by interleaving threads around sync API PPs alone. Here, preempting during the race on line 1 is necessary to force the program into the control flow which can race on line 5.}
	\label{fig:nondet-dr}
\end{figure}

\begin{definition}[Reachable data race]
A data race candidate which will be identified by a MC configured to preempt only on locking API boundaries,
or transitively also configured to use data-race PPs of other (already) reachable data races.
\end{definition}

\begin{definition}[Reachable preemption point]
A PP $\pai$ such that either $\ppinstr{\pai}$ is part of a reachable data race, or a sync API boundary.
\end{definition}

\begin{lemma}
	Given a fully-relevant interleaving
	\[
		\mathcal{I} = \{ ..., (t_{\alpha{}i},\pai), \ppothers{\pai}, (t_{\alpha(i+1)},p_{\alpha(i+1)}), ... \}
	\]
	if $\pai$ is a data-race PP, and the all PPs in the interleaving prefix
	\[
		\mathcal{J} = \{ ..., (t_{\alpha{}i} \cup t_{\alpha(i+1)}',p_{\alpha(i+1)}'), \ppothers(\pai)' \}
	\]
	are reachable, then $\pai$ is reachable.
	\label{lem:coalesce}
\end{lemma}

Note here that in $\mathcal{J}$, some instructions of $t_{\alpha(i+1)}$ may change because they were reordered with $\ppothers(\pai)$;
i.e., $\ppinstr(\pai)$ may not necessarily be the only site of thread communication,
and some other conflicts may cause $t_{\alpha(i+1)}$ to behave differently.
Hence we have labelled it $t_{\alpha(i+1)}'$, and similar for thread $\alpha$'s next PP.
Nevertheless, the first instruction of $t_{\alpha(i+1)}'$ must still be the same as $\ppinstr{\pai}$.

% TODO: What if ppothers changes based on ppinstr(pai)? You say, there must exist some J where ppothers has fewer PPs. You just stop preempting! Those preemptions don't matter. Gotta go fast. Gotta keep the induction metric of getting fewer pps in J than in I.

%%%

\begin{lemma}
	If a bug can be exposed by any fully-relevant interleaving, %some combination of preemptions on any data-race instruction and sync API boundaries,
	Iterative Deepening will eventually test a state space in which all those PPs are enabled\footnote{
		In the paper we noted that when a bug is found, \quicksand~cancels all superset state spaces, anticipating that they would find the same bug.
		Because we are concerned with verifying the absence of bugs, we say this also suffices for convergence, but for brevity, we will leave this condition implicit.}.
	\label{lem:saturation}
\end{lemma}

% TODO: Here's what to do.
% Induct on the PPs by the order of next(my_PP) in the trace.
% Inductive assumption gives that all prior other data race PPs with next(other_PP) before next(my_PP) are reachable,
% Is there a prior data race PP with next(other_PP) *AFTER* next(my_PP)?
% If not, then we are done, and this one is reachable.
% If so, then we ask, which transition among others(other_PP) (this includes us) does instr(other_PP) conflict with?
% <vague handwavy part>
% Then we show that there must be a trace with next(other_PP) interleaved to be before next(my_PP) where, even if the bug doesn't show up, the data race still shows up, which makes other_PP reachable.
% </vague handwavy part>
% Since other_PP is reachable then my_PP is reachable.

\begin{proof}
Let $\mathcal{I} = \{(t_{\alpha{}1}, p_{\alpha{}1}), ...\}$ be the fully-relevant interleaving and PP sequence in the premise of the lemma which exposed a bug.
We proceed by inducting on the PPs according to the order in which for each PP $\pai$, $\ppnext{\pai}$ appeared in the trace.
Note that this is not necessarily the same as the order of the $\pai$s themselves in $\mathcal{I}$;
i.e., it is the order of the racy instructions rather than the order of their immediately preceding preemptions.
%Figure~\ref{fig:why-ppnext} shows why the first PP, ordered by earliest point of preemption, might not necessarily be the first reachable one.
%%% well, the way we do it, might not be the first reachable one either...

\newcommand\pbj{\ensuremath{p_{\beta{}j}}}
For both the base case and inductive step, we know (vacuously, for the former) that for each other PP $\pbj$ with $\ppnext{\pbj}$ occurring before $\ppnext{\pai}$ in $\mathcal{I}$, $\pbj$ is reachable.
% Now to show \pai is reachable.

\begin{itemize}
	\item
	\item
\end{itemize}
\end{proof}

%%%%

\subsection{Conclusion}

\setcounter{theorem}{1}
\begin{theorem}[Convergence of Iterative Deepening]
	If a bug is exposed by an interleaving in the na\"{i}ve state space, Iterative Deepening will eventually test an interleaving which exposes the same bug.
\end{theorem}

\begin{proof}
By Lemma~\ref{lem:equivalent}, there must be some equivalent fully-relevant interleaving.
%the na\"{i}ve interleaving must be equivalent to some interleaving built of data-race PPs and sync API PPs.
By Lemma~\ref{lem:saturation}, Iterative Deepening will eventually discover and enable all necessary data-race PPs, with sync API PPs being enabled by the maximal state space assumption.
Then we simply lean on the soundness of DPOR~\cite{dpor} to guarantee finding the buggy interleaving within this state space.
\end{proof}