### 2015-11-20 pldi-16/proof/di.tex

\section{Definitions}

\subsection{Stateless model checking terms}

\begin{definition}[Preemption point (PP)]
	A PP is a code location between two instructions at which we may force threads to switch.
\end{definition}

In the main paper, we identified PPs by predicates on the instruction pointer and current thread ID. Hence, the ``same'' PP may occur multiple times during an execution; for example, {\tt mutex\_lock()} may be called from {\tt foo()} and later again from {\tt bar()}.
For the sake of this proof, we separate such cases into multiple unique PPs: each PP is simply a label denoting the end of a single transition.

\begin{definition}[Transition]
A sequence of instructions executed by the program between two preemption points (PPs).
\label{def:transition}
\end{definition}
Our model checker provides the invariant that each transition's instructions are associated with exactly one thread. (That is, the set of PPs always includes all thread switches.)
We also include a trace of all memory accesses in the representation of transitions.

\begin{definition}[Must-happen-before (MHB)]
	Given two transitions $A$ and $B$, we say $A$ MHB $B$ if $B$ cannot be reordered to occur before $A$.
\end{definition}
All transitions of the same thread are trivially MHB-related.
Two transitions $A$ and $B$ of different threads MHB if some synchronization event in $A$ causes $B$ to become runnable while it was previously blocked. Such synchronization events include {\tt thread\_create}, {\tt cond\_signal}, {\tt sem\_post}, but {\em not} {\tt mutex\_lock} or {\tt mutex\_unlock}.

Note how our {\em must}-happen-before relation differs from the conventional definition of happens-before (``observed to happen before'') \cite{lamport-clocks}.
Our use of MHB matches the ``limited happens-before'' used in \cite{hybriddatarace} and \cite{tsan};
the advantage of this over pure-happens-before detectors in producing fewer false negatives is well-argued in those prior works\footnote{
Because pure-HB data race detectors avoid false positives altogether, they would have no trouble avoiding our malloc-recycle false positives.
However, as prior work has shown, they miss many other bugs involving unprotected variables accessed alternately before and after mutex-protected critical sections.
Indeed, because most concurrent malloc implementations are protected by a lock,
our malloc-recycle false positives are indistinguishable from such false negatives under pure-HB.
}.
We illustrate the difference in Figure~\ref{fig:mhb}.

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& Thread 1 & Thread 2 \\
	1 & \texttt{\hilight{brickred}{my\_x->foo = ...;}} & \\
	2 & \texttt{\hilight{olivegreen}{mutex\_lock(...);}} &\\
	3 & \texttt{global\_x = my\_x;} & \\
	%4 & \texttt{\hilight{olivegreen}{yield();}} & \\
	4 & \texttt{\hilight{olivegreen}{mutex\_unlock(...);}} & \\
	5 & & \texttt{\hilight{olivegreen}{mutex\_lock(...);}} \\
	6 & & \texttt{my\_x = global\_x;} \\
	7 & & \texttt{\hilight{olivegreen}{mutex\_unlock(...);}} \\
	8 & & \texttt{if (my\_x != NULL)} \\
	9 & & \texttt{\hilight{brickred}{~~~~my\_x->foo = ...;}} \\
\end{tabular}
	\caption{Example program to illustrate the difference between {\em pure happens-before} and {\em must-happen-before}.
	Under pure happens-before (which does not identify false positives), lines 1 and 9 are not a data race candidate.
	Under MHB, they are; although after trying to reorder them, it will be classified as a false positive.}
	\label{fig:mhb}
\end{figure}
Intuitively, $A$ MHB $B$ can be thought to mean ``$B$ cannot be reordered before $A$''.
Note also that, while lock-protected critical sections can be reordered around each other (i.e., line 1 not MHB lines 8-9),
one cannot be reordered to be in the middle of the other (i.e, lines 3-4 MHB line 6).
Hence, MHB is not necessarily transitive.
In the latter case, the MHB relation is established by the mutex's blocking mechanism used during contention.

\begin{definition}[Shared memory conflict]
A pair of memory accesses between two threads to the same address where at least one of them is a write.
\end{definition}

\begin{definition}[Interleaving]
	An ordered list of transitions and preemption points between them.
\end{definition}

\begin{definition}[Independent transitions]
Two transitions between different threads are independent if the intersection of their shared memory accesses contains no conflicts.
\end{definition}

\begin{definition}[Equivalent interleaving]
Two interleavings are equivalent if one can be transformed into the other by permuting only independent transitions.
\end{definition}

Intuitively, the behaviour of a program could change by reordering two transitions only if they contain a memory conflict.
All possible interleavings of a program can be partitioned into equivalence classes,
so only one interleaving from each equivalence class need be tested to ensure total schedule coverage \cite{mazurkiewicz}.

\begin{definition}[State space]
	A set of interleavings subject to the constraint that, given the preemption points used, all equivalence classes of possible interleavings are represented by at least one member.
\end{definition}

\begin{definition}[Dynamic Partial Order Reduction (DPOR)]
	A state-space search algorithm for stateless model checkers,
	guaranteed to reorder transitions of two threads
	iff they are not independent and are not related by MHB \cite{dpor}.
	\label{def:dpor}
\end{definition}

The soundness of DPOR guarantees that if a program behaviour can possibly be exposed by any thread interleaving around the given transitions/PPs,
that interleaving will eventually be tested by reordering only such conflicting transitions.
In other words, reordering memory-independent thread transitions cannot possibly affect program behaviour.

\subsection{Data race and other memory terms}

\begin{definition}[Data race]
A shared memory conflict where furthermore:
\begin{itemize}
	\item The intersection of both threads' locksets is empty (i.e., the threads do not hold the same lock during each access), and
	\item The threads' transitions are not related by MHB.
\end{itemize}
\end{definition}

The same as in the paper, we distinguish between data-race {\em candidates} and data-race {\em bugs}.
In this proof we are concerned solely with candidates, and judging whether they are true data races or false positives.
It is up to \quicksand, outside the scope of this proof, to decide whether true data races are benign or buggy.

\begin{definition}[False positive data race]
	An apparent data race that cannot be observed in the opposite order from what was actually executed.
\end{definition}

False positives are caused when some data dependency based on some other shared state, invisible to the data-race analysis,
changes some variable values when the threads are reordered, such that the memory addresses no longer collide.

\begin{definition}[Malloc-recycle data race]
	A data race where the address is contained in some heap-allocated memory, and between the two accesses, that memory was passed to free() and returned again by a subsequent malloc().
\end{definition}

Figures~\ref{fig:recycle} and \ref{fig:recycle-bug} show an example.
In the case of malloc-recycle false positives, the allocation heap is the ``other shared state'' mentioned in the previous definition, and malloc's return value is the variable value that changed.

Recent work \cite{sparc-ssm} has proposed hardware techniques for detecting many classes of stale heap pointer accesses, including the one shown in Figure \ref{fig:recycle-bug}.
This approach could be combined with stateless model checking in future work to identify such bugs immediately,
rather than requiring Iterative Deepening to explore new state spaces corresponding to the data race.
However, if the {\tt malloc} call were in thread 1 instead of thread 2, the bug would still be nondeterministic, requiring stateless model checking to expose.

\begin{definition}[Use after free]
	Any read or write to heap memory which was once allocated, but no longer is.
\end{definition}

These can immediately be identified as failures by a MC which tracks allocation state.
%Most commonly this refers to accesses to a region already freed, but for brevity we also include
% TODO: need heap block overrun?

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{struct x \{ int foo; int baz; \} *x;}} \\
	& \multicolumn{2}{c}{\texttt{struct y \{ int bar; \} *y;~~~~~~~~~~}} \\
	\\
	& Thread 1 & Thread 2 \\
	1 & \texttt{\hilight{brickred}{x1->foo = ...;}} & \\
	2 & \texttt{\hilight{olivegreen}{free}(x1);} \\
	3 & & \texttt{\hilight{commentblue}{// x's memory recycled}} \\
	4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	5 & & \texttt{\hilight{commentblue}{// ...initialize...}}\\
	6 & & \texttt{publish(y);} \\
	7 & & \texttt{\hilight{brickred}{y->bar = ...;}} \\
\end{tabular}
\caption{False-positive malloc-recycle pattern. This is the common case for which we avoid creating new state spaces.}
\label{fig:recycle}
\end{figure}

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& Thread 1 & Thread 2 \\
	1 & \texttt{publish(x1);} & \\
	2 & & \texttt{x2 = get\_published\_x();} \\
	3 & \texttt{\hilight{brickred}{x1->foo = ...;}} & \\
	4 & \texttt{\hilight{olivegreen}{free}(x1);} \\
	5 & & \texttt{\hilight{commentblue}{// x's memory recycled}} \\
	6 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	7 & & \texttt{\hilight{brickred}{x2->foo = ...;}} \\
\end{tabular}
\caption{Adversarial program which fits the malloc-recycle pattern, but nevertheless contains a true race.}
\label{fig:recycle-bug}
\end{figure}

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& Thread 1 & Thread 2 \\
	1 & \texttt{publish(x1);} & \\
	2 & & \texttt{x2 = get\_published\_x();} \\
	3 & & \texttt{\hilight{commentblue}{// x not free, so malloc's}} \\
	4 & & \texttt{\hilight{commentblue}{// return value changes!}} \\
	5 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	6 & & \texttt{\hilight{brickred}{x2->foo = ...;}} \\
	7 & \texttt{\hilight{brickred}{x1->foo = ...;}} & \\
	8 & \texttt{\hilight{olivegreen}{free}(x1);} \\
\end{tabular}
\caption{Goal interleaving, which reorders the adversarial program's threads away from the pattern, while the data race remains.}
\label{fig:recycle-goal}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Intuition}

In this section we provide (hopefully) intuitive summaries of both our proof goals.
%for readers not interested in double-checking the proofs' internal structure.

{\bf Intuition for Iterative Deepening convergence.}
In summary, we are proving that when Iterative Deepening finishes saturating the set of available data-race PPs,
that set represents every single code location where a preemption could possibly affect the program's behaviour,
and that completing the associated state spaces is as strong of a verification as testing all possible thread interleavings under any preemptions anywhere.
Some data-races may be hidden in control-flow paths which could only be executed after preempting on a different data-race,
but the technique's iterative nature will eventually find it.
On the other hand, relying on the soundness of DPOR, preempting on an instruction which is neither a data-race or sync API boundary cannot affect the program's behaviour,
so any PPs beyond the ones we consider are irrelevant.

{\bf Intuition for malloc-recycle soundness.}
In summary, we are proving that if a malloc-recycle-pattern data race is a true race, rather than a false positive,
then DPOR is guaranteed to ``reorder away the free and re-malloc''.
In other words, DPOR's exploration will eventually interleave threads in such a way that the malloc-recycle pattern will disappear,
while the access pair remains for the data-race detector to find, as shown in Figure~\ref{fig:recycle-goal}.
Hence, in the same state space where the malloc-recycle data race was found, if it's a true race, the same race will also appear without the recycle pattern.
So if that race hides a failure bug (assertion, segfault, etc.), Iterative Deepening will still be led to the necessary preemption point to find that bug.