### 2016-05-20 oopsla-16/response.txt
add prelim author response

We agree with reviewer A's concern about possible bias in the set of
benchmarks. To address this, we will add some analysis of the types of bugs
found. For example, in the Iterative Context Bounding control experiment, 82
bugs were found with preemption bound 1, 15 with bound 2, and 2 with bound 3.
This is consistent with the distribution in [1] and shows our bugs are not
biased towards requiring high preemption bounds.

Our focus on maximizing completed state spaces is motivated by [2], in which
developers prefer compositional testing (manually restricting a test's scope
to complete faster), which suggests that full completions of partial state
spaces are easier for humans to interpret than partial completions of full
spaces. Our evaluation also shows that subset state spaces generally uncover
bugs faster (QS-Sync-Only vs SSS-MC-DPOR). However, your suggestion to
sometimes complete larger jobs first, thereby pruning any subsumed smaller
jobs, is an insightful optimization which we could implement in future work.

If the race detector produces a candidate where only one of two orders is
feasible, the MC will subsequently classify it as a false positive.

To address reviewer B, we choose new preemption points to add for each
data-race candidate, one for each racing instruction. This produces several
new jobs, each with a different combination of the new ones with existing
ones. For the sake of soundness, which assumes an infinite CPU budget, it
would suffice to randomly choose among new jobs until all are exhausted, or
even to always choose the biggest (maximal) job and ignore subset jobs.
Section 5's proofs assume this simple approach (call it Algorithm 0).
Algorithm 1 in section 3 optimizes algorithm 0 for finite CPU budgets to
prioritize finding bugs faster. Sorry for the confusion and we will clarify
this, including Algorithm 0 explicitly.

Regarding the suggestion of Maximal Causality Reduction [3], we were not able
to do a direct comparison because that paper has no open-source
implementation. We agree it is not necessarily orthogonal to our approach, but
could be combined in future work when an implementation becomes available.

We will clarify in our introduction that we assume sequentially-consistent
hardware, although in principle our work could be extended to weak memory
models using the techniques in [4].

Regarding Figure 3, the contrived program is just an example to explain the
background. We agree other approaches exist which can solve this case.

Reviewer C insightfully observes that Figure 7's count of bugs excludes
data-race candidates from the control experiments. This is the mentioned
philosophical difference about data races versus concrete failures; to address
this, we will add a table listing how many data races the control experiments
reported, and how many of those Quicksand refuted as false positives.

Reviewer D mentions that both HB and Limited HB will find all data races when
combined with model checking. We suspected this was true, but were only able
to prove the case for Limited HB. If you could provide a citation which proves
the case of regular HB combined with model checking, we are interested to read
it and would gladly include it for cam-ready. However, one further advantage
of Limited HB is that it reports candidates sooner, allowing Quicksand to test
them immediately, at the expense of more false positives. We will add a
discussion of this trade-off.

Regarding our malloc-recycle solution, we tried your suggestion to never
reallocate memory, but this unacceptably impacted the performance of some
malloc-heavy tests. Our proof allows eliminating these data-race candidates
without hacking the allocator.

Thanks for the clarification about FastTrack and CHESS; we will improve our
discussion of those tools.

Finally, reviewers A and D both criticized the formalism for unclear
definitions and notation. We agree and will work to improve the foundation and
notation definitions for cam-ready. For example, in the definition of MHB, we
define "cannot be reordered" in terms of the ability for threads to be
scheduled at prior preemption points, with no intermediate blocking operation
by the latter thread (this coincides with DPOR's approach to identify
alternate interleavings to explore).

[1] Musuvathi and Qadeer, Iterative Context Bounding, PLDI07
[2] Ball et al, Preemption Sealing, TACAS10
[3] Huang, Maximal Causality Reduction, PLDI15 ([24] in the paper/reviews)
[4] Zhang et al, DPOR for relaxed memory, PLDI15