### 2016-07-19 oopsla-16/coverletter/letter.tex
add cover letter skeleton

\documentclass{article}
\special{papersize=8.5in,11in}

%\usepackage[charter]{mathdesign}
\usepackage{fullpage}

\title{OOPSLA Round 2 Cover Letter \\ {\large Paper \#177: Stateless Model Checking with Data-Race Preemption Points}}
\author{Ben Blum (\textsf{bblum@cs.cmu.edu}) \and Garth Gibson (\textsf{garth@cs.cmu.edu})}
\date{July 19, 2016}

\begin{document}

\maketitle
\thispagestyle{empty}

\noindent To the OOPSLA Program Committee,
\\

Thank you for your helpful reviews. Following is a list of revisions we have made, sorted by which of the mandatory revisions they address.

\begin{enumerate}
	\item {\bf Acknowledge sequentially-consistent memory assumption.}
	\begin{itemize}
		\item ...
	\end{itemize}
	\item {\bf Implement and compare with precise Happens-Before.}
	\begin{itemize}
		\item ...
	\end{itemize}
	\item {\bf Explain how we add new preemption points.}
	\begin{itemize}
		\item ...
	\end{itemize}
	\item {\bf Elaborate on related work (MCR and FastTrack).}
	\begin{itemize}
		\item ...
	\end{itemize}
	\item {\bf Philosophical difference between data races and failures.}
	\begin{itemize}
		\item ...
	\end{itemize}
\end{enumerate}

We have also made some changes that were not included in the mandatory revisions, to address other reviewer comments and in one case to fix some bugs in our experiments, as follows:

\begin{itemize}
	\item ...
\end{itemize}

\noindent Regards,
\\

\noindent Ben Blum \\
Garth Gibson

\end{document}