### 2016-08-16 sigcse-17/paper.tex

\documentclass{sig-alternate-05-2015}

\usepackage{xspace}
\usepackage{color}
\usepackage{graphicx}
\usepackage{listings}

\lstset{language=C,
        frame=none,%single,
        basicstyle=\ttfamily,
        keywordstyle=\bfseries,
        commentstyle=\em,
        tabsize=2,
        showstringspaces=false,
        flexiblecolumns=false,
        mathescape=true,
        escapeinside={\#}{\#},
        % numbers=left, numberstyle=\tiny, numbersep=5pt %, stepnumber=2, 
        numbersep=5pt,
        % literate={->}{{$\rightarrow$}}1 {<->}{{$\leftarrow$}}1
}
\newcommand\x[1]{\lstinline{#1}}

\begin{document}

\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% TODO: ???
%\doi{10.475/123_4}
%\isbn{123-4567-24-567/08/06}
%\acmPrice{\$15.00}
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.

\conferenceinfo{SIGCSE '17}{March 8th - 11th, Seattle, Washington, USA}
\CopyrightYear{2017}

% TODO camready
\newcommand\carnegiemellon{{\em Double-Blind University}\xspace}
\newcommand\cmu{DBU\xspace}
\newcommand\pebbles{\cmu-OS\xspace}
\newcommand\landslide{Landslide\xspace}
\newcommand\fourten{CS999}
%\newcommand\carnegiemellon{{\em Carnegie Mellon University}\xspace}
%\newcommand\cmu{CMU\xspace}
%\newcommand\pebbles{Pebbles\xspace}
%\newcommand\landslide{Landslide\xspace}
%\newcommand\fourten{15-410}

\title{Teaching Low-Level Concurrency with \pebbles and \landslide}

\numberofauthors{3}

% TODO camready
\author{
	\em double-blind submission
%	\alignauthor
%	Ben Blum \\
%	\affaddr{Carnegie Mellon University} \\
%	\affaddr{5000 Forbes Avenue} \\
%	\affaddr{Pittsburgh, PA, USA} \\
%	\email{bblum@cs.cmu.edu}
%	\alignauthor
%	David A. Eckhardt \\
%	\affaddr{Carnegie Mellon University} \\
%	\affaddr{5000 Forbes Avenue} \\
%	\affaddr{Pittsburgh, PA, USA} \\
%	\email{davide@cs.cmu.edu}
%	\alignauthor
%	Garth Gibson \\
%	\affaddr{Carnegie Mellon University} \\
%	\affaddr{5000 Forbes Avenue} \\
%	\affaddr{Pittsburgh, PA, USA} \\
%	\email{garth@cs.cmu.edu}
}

\maketitle
\begin{abstract}

In \carnegiemellon's undergraduate operating systems course
students learn how to write correct and principled concurrent code through
a series of low-level programming projects.
The curriculum includes implementing hardware device drivers and a user-level
threading library, and culminates in a six-week long project in which students
build from the ground up a fully-preemptible, UNIX-like kernel that can run
on real hardware.

We aspire to expand the scope of these projects in response to
typical hardware platforms'
increasing complexity,
especially massive multi-core.
To that end, we wish to reduce the time students spend
debugging concurrency bugs.
We present Landslide, a
%systematic testing
stateless model checker which
offers a more principled approach than stress testing.
% edupar version
%We report on our preliminary experience working with students using Landslide to find previously-overlooked bugs in their own code, and discuss future methods for making Landslide more accessible to struggling students.
% sigcse version
We report on our experience giving students Landslide during the thread library
%and kernel projects.
project,
%Although kernel-mode Landslide requires manual instrumentation, testing thread libraries is fully automatic.
% TODO: verify these, obv. ;)
showing that it allows them to find and fix previously-overlooked bugs.
Further, %even struggling students can take advantage of it,
Landslide's users achieve higher concurrency scores on average in subsequent projects,
% FIXME awkward phrasing
suggesting that debugging with a model checker produces better student retention of concurrency principles.

\end{abstract}

% TODO, do dis
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

% End generated code

\printccsdesc

\keywords{concurrency, debugging, model checking, computer science education}

\input{intro}

\section{The {\secit Body} of The Paper}
% TODO
TODO body

\section{Conclusions}
% TODO
TODO conclusion

% TODO camready
% \section{Acknowledgments}
% Many people have contributed to the development of
% 15-410 over the past decade.
% Stephen Muckle, currently at Qualcomm,
% led the switch to x86 projects running in Simics.
% Nathaniel Wesley Filardo,
% currently a Ph.D.\ candidate at Johns Hopkins,
% % increased the sanity of the build infrastructure,
% designed the \x{swexn()} system call and helped critique this paper.
% The current reference kernel, called ``pathos,''
% was written by Michael J. Sullivan,
% currently a CMU CS Ph.D.\ candidate,
% and Elly Fong-Jones, currently at Google.
% The SMP infrastructure code was contributed by
% Ryan Pearl, currently at Dropbox.
% Joshua Wise, currently at NVIDIA,
% led several kernel-extension projects.
% % including the hypervisor project.
% The course has benefitted
% greatly from free educational Simics licenses
% donated by Virtutech/Wind River.
% Code-size numbers were generated using David A.\ Wheeler's
% ``SLOCCount.''
% Work on Landslide was sponsored by the U.S. Army Research Office under grant number W911NF0910273.

\bibliographystyle{abbrv}
\bibliography{citations}

%\balancecolumns
% That's all folks!
\end{document}