### 2016-08-19 sigcse-17/landslide.tex
sigcse: write about landlside

\section{Landslide}
\label{sec:landslide}

Motivated by the stress tests' inconsistency at finding concurrency bugs,
% TODO camready
we have built Landslide%\cite{landslide}
, a stateless model checker for \pebbles kernels and thread libraries.
Stateless model checking \cite{verisoft} offers an alternative
%concurrency testing approach
to stress testing,
in which the testing framework controls scheduling to force a new thread interleaving to execute in each iteration of a test.
It follows in the footsteps of related tools such as CHESS \cite{chess},
but also provides several important features for use in an educational setting,
such as heuristics to handle unusual student solutions, kernel-level testing, and user-friendly progress reports and HTML debugging traces.

\subsection{Design}

Conceptually,
Landslide is designed as follows.
Through several annotations in the kernel source, Landslide tracks
%the scheduler state to know
which threads are running or runnable.
(For testing thread libraries, these annotations are already provided in the reference kernel binary;
for testing kernels, the student must add them by hand.)
When execution reaches certain interesting {\em preemption points},
Landslide triggers artificial clock interrupts
to force the scheduler to run a different thread.
When a test finishes execution according
to one pattern of thread switches,
Landslide rewinds the simulation state
and resumes the test according to a different interleaving.
After each instruction,
Landslide applies several bug-detection predicates to
the program state to detect
illegal heap accesses,
deadlock, infinite loops, and panics.

{\bf Example.}
Figure \ref{fig:paradise}(a) shows an example concurrency bug which we will use as a running example.
The code implements a producer/consumer program, in which one thread sends data to the other through a shared list,
synchronized with a mutex and condition variable.
However, because the consumer code only checks the list size once before blocking,
multiple simultaneous consumers can race,
causing a segfault when one receives a null \x{data} pointer. %and tries to dereference it

{\bf State space exploration.}
To trigger the interleaving which exposes this bug, threads must be preempted at the boundaries of the synchronization primitives (either when \x{cond_wait} blocks or at the end of \x{mutex_unlock}).
Landslide automatically considers these boundaries to be {\em preemption points}: candidate sites for timer-driven thread switches.
It can also add preemption points dynamically,
as identified by an integrated data-race analysis \cite{tsan,fasttrack},
to expose shared-memory bugs when synchronization is absent.

The branching possibilities arising from each preemption point produce an execution tree, or state space,
as shown in Figure \ref{fig:paradise}(b).
%As these state spaces are exponentially-sized,
Landslide uses Dynamic Partial Order Reduction \cite{dpor} to identify and prune equivalent interleavings
to mitigate the exponential explosion of these state spaces.
% TODO make sure this gets mentioned later, if not here
%and state space estimation \cite{estimation} to predict which
% TODO CAMREADY: cite oopsla
% In principle, the combination of synchronization API and data-race preemption points is sufficient to expose all bugs that are reachable by any combination of timer interrupts on any instructions.

\subsection{Identifying Bugs}

Landslide checks for bugs in several ways.
It can identify assertion failures, use-after-free and other invalid heap accesses, and
deadlocks, and also includes a heuristic check for infinite loops and livelock.
%Note that our use of the term ``race condition'' throughout this paper refers to concurrency errors in general, a category that includes data races, atomicity violations, and nondeterministic deadlocks.
%In contrast with analyses that directly detect data races~\cite{tsan},
Although Landslide employs a data-race analysis to decide where to preempt,
it does not directly classify suspicious memory accesses as bugs,
but rather seeks concrete, observable failures that can arise from many different types of errors.

% TODO: make fig trace

When Landslide identifies a bug, it outputs an HTML-formatted
{\em preemption trace}.
Figure~\ref{fig:trace} shows a typical output for the example bug in Figure \ref{fig:paradise}.
The trace reports what kind of bug was detected
and each preemption point in the current interleaving:
which thread was running, a trace of its stack when it was preempted, and the thread that ran next.
Hence, students can bypass tediously reconstructing the schedule using \x{printf},
and hopefully better understand the concurrent execution that exposed the bug,

\subsection{Heuristics}

TODO talk about heuristics