### 2016-08-27 sigcse-17/paper.tex

\documentclass{sig-alternate-05-2015}

\usepackage{xspace}
\usepackage{color}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{enumitem}
%\usepackage{colortbl}

\lstset{language=C,
        frame=none,%single,
        basicstyle=\ttfamily,
        keywordstyle=\bfseries,
        commentstyle=\em,
        tabsize=2,
        showstringspaces=false,
        flexiblecolumns=false,
        mathescape=true,
        escapeinside={\#}{\#},
        % numbers=left, numberstyle=\tiny, numbersep=5pt %, stepnumber=2,
        numbersep=5pt,
        % literate={->}{{$\rightarrow$}}1 {<->}{{$\leftarrow$}}1
}
\newcommand\x[1]{\lstinline{#1}}

\newcommand{\sect}[1]{\S #1}

\setlist[itemize]{leftmargin=*}

\begin{document}

\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% TODO: ???
%\doi{10.475/123_4}
%\isbn{123-4567-24-567/08/06}
%\acmPrice{\$15.00}
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.

\conferenceinfo{SIGCSE '17}{March 8th - 11th, Seattle, Washington, USA}
\CopyrightYear{2017}

% TODO camready
\newcommand\carnegiemellon{{\em Double-Blind University}\xspace}
\newcommand\cmu{DBU\xspace}
\newcommand\pebbles{EthOS\xspace}
\newcommand\landslide{Logos\xspace}
\newcommand\fourten{CS999\xspace}
\newcommand\twothirteen{CS888\xspace}
%\newcommand\carnegiemellon{{\em Carnegie Mellon University}\xspace}
%\newcommand\cmu{CMU\xspace}
%\newcommand\pebbles{Pebbles\xspace}
%\newcommand\landslide{Landslide\xspace}
%\newcommand\fourten{15-410}

\title{Teaching Low-Level Concurrency with \pebbles and \landslide}

\numberofauthors{3}

% TODO camready
\author{
	\em double-blind submission
%	\alignauthor
%	Ben Blum \\
%	\affaddr{Carnegie Mellon University} \\
%	\affaddr{5000 Forbes Avenue} \\
%	\affaddr{Pittsburgh, PA, USA} \\
%	\email{bblum@cs.cmu.edu}
%	\alignauthor
%	David A. Eckhardt \\
%	\affaddr{Carnegie Mellon University} \\
%	\affaddr{5000 Forbes Avenue} \\
%	\affaddr{Pittsburgh, PA, USA} \\
%	\email{davide@cs.cmu.edu}
%	\alignauthor
%	Garth Gibson \\
%	\affaddr{Carnegie Mellon University} \\
%	\affaddr{5000 Forbes Avenue} \\
%	\affaddr{Pittsburgh, PA, USA} \\
%	\email{garth@cs.cmu.edu}
}

\maketitle
\begin{abstract}

In \carnegiemellon's undergraduate operating systems course
students learn how to write correct and principled concurrent code through
a series of low-level programming projects.
The curriculum includes implementing hardware device drivers and a user-level
threading library, and culminates in a six-week long project in which students
build from the ground up a fully-preemptible, streamlined UNIX-like kernel that can run
on real hardware.

We aspire to expand the scope of these projects in response to
typical hardware platforms'
increasing complexity,
especially massive multi-core.
To that end, we wish to reduce the time students spend
debugging concurrency bugs.
We present \landslide, a
%systematic testing
stateless model checker which
offers a more principled approach than stress testing.
% edupar version
%We report on our preliminary experience working with students using \landslide to find previously-overlooked bugs in their own code, and discuss future methods for making \landslide more accessible to struggling students.
% sigcse version
We report on our experience giving students \landslide during the thread library
%and kernel projects.
project,
%Although kernel-mode \landslide requires manual instrumentation, testing thread libraries is fully automatic.
showing that it allows them to find and fix previously-overlooked bugs.
%% no :( :(
% Further, %even struggling students can take advantage of it,
% \landslide's users achieve higher concurrency scores on average in subsequent projects,
% suggesting that debugging with a model checker improves student retention of concurrency principles.

\end{abstract}

% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
\begin{CCSXML}
	 <ccs2012>
	 <concept>
	 <concept_id>10011007.10011006.10011073</concept_id>
	 <concept_desc>Software and its engineering~Software maintenance tools</concept_desc>
	 <concept_significance>500</concept_significance>
	 </concept>
	 <concept>
	 <concept_id>10010147.10011777.10011778</concept_id>
	 <concept_desc>Computing methodologies~Concurrent algorithms</concept_desc>
	 <concept_significance>300</concept_significance>
	 </concept>
	 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Software maintenance tools}
\ccsdesc[300]{Computing methodologies~Concurrent algorithms}

% End generated code

\printccsdesc

\keywords{concurrency, debugging, model checking, computer science education}

\input{intro}
\input{410}
\input{landslide}
\input{eval}

\section{Conclusions and Future Work}

We are presently extending \landslide to support Pintos kernels \cite{pintos},
%used at many top CS schools including Stanford, Berkeley, and University of Chica
another instructional kernel popular at many top CS schools.
%a more widely-used instructional kernel than \pebbles.
Because Pintos mandates more internal APIs than \pebbles,
\landslide can more easily instrument student code without requiring manual annotations.
In a preliminary experiment, we tested 79 student Pintos implementations with \landslide,
finding 59 concurrency bugs with 10 CPU-hours allocated to each of 3 test cases \cite{quicksand-anonymized}.
In future work, we hope to fully automate Pintos instrumentation, and offer \landslide at other schools beyond \cmu.

%\fourten, the undergraduate operating systems course at \cmu, teaches concurrency in a

% TODO camready
% \section{Acknowledgments}
% Many people have contributed to the development of
% 15-410 over the past decade.
% Stephen Muckle, currently at Qualcomm,
% led the switch to x86 projects running in Simics.
% Nathaniel Wesley Filardo,
% currently a Ph.D.\ candidate at Johns Hopkins,
% % increased the sanity of the build infrastructure,
% designed the \x{swexn()} system call and helped critique this paper.
% The current reference kernel, called ``pathos,''
% was written by Michael J. Sullivan,
% currently a CMU CS Ph.D.\ candidate,
% and Elly Fong-Jones, currently at Google.
% The SMP infrastructure code was contributed by
% Ryan Pearl, currently at Dropbox.
% Joshua Wise, currently at NVIDIA,
% led several kernel-extension projects.
% % including the hypervisor project.
% The course has benefitted
% greatly from free educational Simics licenses
% donated by Virtutech/Wind River.
% Code-size numbers were generated using David A.\ Wheeler's
% ``SLOCCount.''
% Work on \landslide was sponsored by the U.S. Army Research Office under grant number W911NF0910273.

\bibliographystyle{abbrv}
\bibliography{citations}

%\balancecolumns
% That's all folks!
\end{document}