### 2016-09-01 oopsla-16/paper.tex

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass[10pt,times,numbers]{sigplanconf}
\usepackage[final]{oopsla2016}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% numbers       To obtain numeric citation style instead of author/year.

\usepackage{amsmath}
\usepackage{url}
\usepackage{enumitem}
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{amsthm}
\usepackage{flushend}
\usepackage{color}
\usepackage[colorlinks=false,breaklinks,draft=false]{hyperref}
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}

\begin{document}

\newcommand\landslide{\textsc{Landslide}}
\newcommand\quicksand{\textsc{Quicksand}}
\newcommand\simics{\textsc{Simics}}
\newcommand{\sect}[1]{\S #1} 
\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{orange}{RGB}{192,96,0}
\definecolor{olivegreen}{RGB}{0,127,0}
\definecolor{brickred}{RGB}{192,0,0}
\definecolor{commentblue}{RGB}{0,0,192}

\newcommand\numthrlibs{79}
\newcommand\numpintoses{78}
\newcommand\numstudence{157} % total pintoses plus p2s

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{OOPSLA '16}{November 2--4, 2016, Amsterdam, the Netherlands}
\copyrightyear{2016}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
\copyrightdoi{nnnnnnn.nnnnnnn}

% Uncomment the publication rights you want to use.
%\publicationrights{transferred}
\publicationrights{licensed}     % this is the default
%\publicationrights{author-pays}

%\titlebanner{banner above paper title}        % These are ignored unless
%\preprintfooter{short description of paper}   % 'preprint' option specified.

% presubmit revisions
\newcommand\revision[1]{#1}
\definecolor{purple}{RGB}{192,0,192}
%\newcommand\revision[1]{\hilight{purple}{#1}}
% 2nd round revisions
%\newcommand\revisionx[1]{\hilight{purple}{#1}\xspace}
\newcommand\revisionx[1]{#1}
% cam ready revisions
%\newcommand\revisionz[1]{\hilight{purple}{#1}\xspace}
\newcommand\revisionz[1]{#1\xspace}

% obfuscate title just a little bit to avoid the billion inevitable google searches
% i don't mind if really dedicated people find their way here, but not everyone please
\newcommand\titlec{l Check}
\newcommand\titleg{ion Points}
\newcommand\titleb{less Mode}
\newcommand\titlee{h Data-R}
\newcommand\titled{ing wit}
\newcommand\titlef{ace Preempt}
\newcommand\titlea{State}

\title{\titlea\titleb\titlec\titled\titlee\titlef\titleg}

\authorinfo{Ben Blum}{Carnegie Mellon University, USA}{bblum@cs.cmu.edu}
\authorinfo{Garth Gibson}{Carnegie Mellon University, USA}{garth@cs.cmu.edu}

\maketitle

\begin{abstract}
Stateless model checking is a powerful technique for testing concurrent programs,
but suffers from exponential state space explosion when the test input parameters are too large.
Several reduction techniques can mitigate this explosion,
but even after pruning equivalent interleavings, the state space size
%for a fixed set of preemption points
is often intractable.
Most prior tools are limited to preempting only on synchronization APIs,
which reduces the space further, but can miss unsynchronized thread communication bugs.
%
Data race detection, another concurrency testing approach, focuses on suspicious memory access pairs during a single test execution.
It avoids concerns of state space size,
%but is prone to false positives:
%spurious reports of access pairs that cannot be reordered to produce a bug.
%but data races do not always correspond to observable failures.
but may report races that do not lead to observable failures,
which jeopardizes a user's willingness to use the analysis.

\revision{We present \quicksand, a new stateless model checking framework}~which
manages the exploration of many state spaces using different preemption points.
It uses state space estimation to prioritize jobs most likely to complete in a fixed CPU budget,
and it incorporates data-race analysis to add new preemption points on the fly.
Preempting threads during a data race's instructions
%These preemption points allow us to check whether each data-race can produce an observable failure, and
can automatically classify the race as buggy or benign,
and uncovers new bugs not reachable by prior model checkers.
%
It also enables full verification of all possible schedules when every data race is verified as benign within the CPU budget.
In our evaluation, \quicksand~found
%with data-race preemption points
\revisionx{1.25x} as many bugs and verified 4.3x as many tests compared to \revision{prior model checking approaches}. %stateless model checking without data-race preemption points.
%Our evaluation shows this technique is
%more effective than single-state-space model checking
%%both at finding more bugs and at completing more state spaces when no bug exists.
%both at finding the same bugs faster and at finding new bugs entirely.
\end{abstract}

\revision{\category{D.2.4}{Software Engineering}{Software/Program Verification}}
\keywords{model checking, data races, verification}
%\category{D.2.5}{Software Engineering}{Testing and Debugging}
%
%% general terms are not compulsory anymore,
%% you may leave them out
%\terms
%term1, term2
%
%\keywords
%keyword1, keyword2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{intro}
\input{overview}
\input{design}
\input{implementation}
\input{soundness}
\input{evaluation}
\input{warpzone}
\input{related}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}
\label{sec:conclusion}

%We are great. Accept our paper.

We have presented Iterative Deepening and \quicksand, a new technique and tool for automating the choice of preemption points (PPs) during stateless model checking.
\quicksand~incorporates data-race analysis to create new PPs tailored specifically to the program under test,
and automatically finds state spaces that are appropriately sized to complete in a given CPU budget.

We achieve better bug-finding results than either single-pass data-race detection or single-state-space model checking alone,
finding new bugs with data-race PPs that could not be exposed by preempting only on synchronization APIs.
Moreover, when all data-race PPs can be fully tested within the CPU budget, we provide a verification as strong as if every single instruction had been used as a PP.
Between data-race bugs and verifications, \quicksand~is shown to provide the best of both worlds.
By using \numstudence~student operating system implementation projects as our test suite, we also show the potential benefit as a debugging platform in educational settings.

%Although \landslide~itself is not publicly available due to its dependency on the non-free simulator \simics, \quicksand~is open-source and its interface can be adapted to fit any similar tool.
\quicksand~is open-source and its interface can be adapted to fit any tool similar to \landslide.
We have also posted our evaluation's data set. %and logs.
They are available at:

\url{https://github.com/bblum/oopsla-quicksand}

% TODO CAMREADY: Make data set, at least, public.
\url{https://github.com/bblum/oopsla-dataset}

% TODO CAMREADY: Don't forget to anonymize the names of all your grupos.

\acks

Many thanks to Vaishaal Shankar, Haryadi Gunawi, and David A. Eckhardt for generously providing student implementations from Berkeley's, U. of Chicago's, and CMU's OS classes respectively.
Thanks to Wind River for the use of their simulator \simics.
Thanks to
Ji\v{r}\'{i} \v{S}im\v{s}a, Joshua Wise, Michael J. Sullivan, Carlo Angiuli, Jean Yang, Brandon Lucia, Haryadi Gunawi, John Wilkes, and
the anonymous OOPSLA reviewers for their helpful comments.
%
We also thank the member companies of the PDL Consortium (Broadcom,
Citadel, EMC, Facebook, Google, Hewlett-Packard Labs, Hitachi, Intel,
Microsoft Research, MongoDB, NetApp, Oracle, Samsung America, Seagate,
Tintri, Two Sigma, Uber, Veritas and Western Digital).
%
This work was supported in part by
the U.S. Army Research Office under grant number W911NF0910273.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}
\bibliography{citations}{}

\end{document}