### 2016-09-01 oopsla-16/proof/assumptions.tex
oopsla proofs: brevity and clarity in intro material

\section{Assumptions}

This section documents our assumptions about the concurrency model, language model, and test environment,
and discusses some limitations that may arise therefrom.

\subsection{Assumptions for both proofs}

{\bf Maximal state space.}
We assume the model checker has all
%static/hard-coded
synchronization API
PPs enabled during this state space exploration,
and that we are not limited by a pressing CPU budget.
%Further, we assume that the static PPs include all lock/unlock/trylock operations on mutexes (or whatever other low-level locks are used) and also all higher-level sync primitives which can cause HB (either directly, or because they are built on top of mutexes).
We assume that all synchronization primitives not listed in our system model are built upon them,
or otherwise annotated for the MC to add additional PPs for them.

Using only a subset of PPs or aborting early due to time-out could each ruin our ability to reach the goal interleaving or goal state space.
However, Iterative Deepening aims to test the most important interleavings with the time available,
so in the case of not enough time,
%our point here is that
continuing the current state space fits that goal best.
%An alternate approach would be to accept the malloc-recycle PPs, giving them the lowest possible priority, just in case any CPUs would otherwise be idle with not enough jobs to run.
%In that case, our point is to justify deprioritizing those jobs so aggressively.

%{\bf Low overhead.}
%We assume the model checker can identify malloc-recycle data races with little or no overhead beyond what's already associated with data race detection.
%Our MC already tracks the heap state, so we implement this check for free with a simple generation counter.

{\bf Shared memory thread communication.}
We assume that the only way for two threads' transitions to affect each other's behaviour, should they be reordered,
is through either shared memory or a correctly-instrumented sync API.
Both DPOR and data-race detection rely on this assumption,
as any other way for threads to affect each other's behaviour could invisibly reduce independence and break soundness.
For brevity in these proofs, we refer to all thread communication as shared memory,
and assume that other mechanisms, such as system calls that access the filesystem, could be instrumented to fit the same model.

{\bf Schedule nondeterminism only.}
We discount the possibility of other types of nondeterminism,
such as program input nondeterminism (including randomness/timestamps) or
store-buffer nondeterminism on weak-memory architectures.
We refer the reader to \cite{klee,portend} for related work on the former, and to \cite{tsopso} for the latter.

\subsection{Assumptions for Iterative Deepening convergence}

{\bf Locks are correct.}
Because hybrid data-race detection uses lockset analysis to exclude many candidate access pairs,
%conclude that many access pairs could not possibly race,
we assume that no preemption during a lock-protected critical section could cause a contending thread to make progress.
This extends to use of disabling/enabling interrupts in kernel code, which we model as a single global lock,
although we do not model RCU \cite{rcu}.

Should the user wish to verify these properties of locks,
they could either run a locking test separately (as we suggest in our paper),
which would cheaply test the locks to an extent limited by the separate test case;
or they could remove the data-race analysis's lockset tracking,
which would expensively test all the main program's required locking properties in tandem with the program itself.

We also assume that any unusual locking discipline, such as recursive mutexes or lock hand-off,
is correctly annotated to the data-race analysis.
Our implementation also models the behaviour of r/w locks and 1-initialized semaphores (the latter heuristically),
although this is tangential to the proof.

%{\bf No forcibly blocking other threads.}
%We assume the sync API behaves much like pthreads and/or message-passing:
%that is, a thread can cause itself to block on some condition,
%and other threads may wake it up,
%but there exists no primitive which one thread can use to revoke another thread's runnable status when that other thread is not itself using the sync API concurrently.
%
%We require this because of the asymmetry between synchronization PPs (which occur {\em after}) synchronization operations)
%and data-race PPs (which occur {\em before} their associated read/write).
%For example, if a primitive $\mathsf{deschedule\_other}$ existed,
%a transition such as $\mathsf{write}(a,v); \mathsf{deschedule\_other}(t_2)$ would lack a PP between the two instructions,
%which could be necessary to expose some...

{\bf Halting.}
From a certain perspective, convergence is the same as completeness,
which should be impossible for any runtime analysis of a Turing-complete language \cite{entscheidungsproblem}.
However, being already limited by practical real-world CPU budgets,
we are already accepting that many tests will time out.
We are concerned with the limited case of when \quicksand~{\em does} terminate with all state spaces completed.
%, the verification is sound.

Inversely, when we prove that Iterative Deepening will eventually find an arbitrary buggy interleaving,
we assume that no intermediate state spaces contain nontermination bugs.
If they do,
%for the sake of testing/verification,
we are satisfied with finding that bug instead.
In this proof, we assume that the MC's heuristic infinite loop checker has no false negatives;
i.e., it will never get stuck forever in an infinite loop without identifying a bug (necessarily accepting some false positives).

\subsection{Assumptions for malloc-recycle soundness}

{\bf Malloc is a magic black box.}
We assume the malloc implementation is correct (e.g., it won't double-allocate),
although we don't assume any implementation details such as a tendency to reuse blocks or allocate adjacent ones.
%particular allocation pattern regarding adjacency/coalescing/reuse.
In fact,
%in our experiments we instruct our MC to
our implementation ignores all potential PPs arising from malloc's internal mutex;
in essence treating it as a ``magic primitive'', because we are not interested in verifying its implementation.
Furthermore, we configured DPOR to ignore any shared memory conflicts arising from internal heap metadata
%when tracking shared memory conflicts
to achieve greater state space reduction.
(If the only consequence of reordering two transitions is malloc returning different addresses, we consider them independent).
%for the purpose of avoiding malloc-only shared memory conflicts.
This is not without consequences; see section~\ref{sec:owned}.

{\bf Sharing heap addresses.}
Finally, we assume that the only way the program can obtain heap addresses is through the return value of $\mathsf{malloc}$.
Because we are testing C programs, any bizarre violations of this assumption are technically possible,
but should you wish to check for bugs like this,
%we would recommend a data-flow analysis which is much cheaper than model checking anyway.
symbolic execution \cite{klee} would be more appropriate.

In Section \ref{sec:proof}, we further assume a $\mathsf{malloc}$ed block's address cannot be obtained through arithmetic on the address of a different block;
in Section \ref{sec:owned} we account for this case by relaxing the previous ``black box'' assumption.