### 2016-09-06 oopsla-16/proof/proof.tex
Finish oopsla proofs for TRification

%
% LaTeX template for prepartion of submissions to PLDI'15
%
% Requires temporary version of sigplanconf style file provided on
% PLDI'15 web site.
%
%\documentclass[pldi]{../sigplanconf-pldi15}
\documentclass[10pt,times,numbers]{../sigplanconf}
\usepackage[proof]{../oopsla2016}

%
% the following standard packages may be helpful, but are not required
%
\usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{amsthm}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{cmu-techreport}
\usepackage{flushend}
%\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI
%\doublespacing

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}

\begin{document}

%
% any author declaration will be ignored  when using 'plid' option (for double blind review)
%

%\newcommand\landslide{\textsc{Landslide}} % Don't mention landslide in the proof.
\newcommand\quicksand{\textsc{Quicksand}}
\newcommand\simics{\textsc{Simics}}
\newcommand{\sect}[1]{\S #1}
\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{olivegreen}{RGB}{0,127,0}
\definecolor{brickred}{RGB}{192,0,0}
\definecolor{commentblue}{RGB}{0,0,255}

%\title{Soundness Proof for Eliminating Malloc-Recycle Data Races in Stateless Model Checking}
\title{\bf Soundness Proofs for Iterative Deepening}
%\authorinfo{Ben Blum}{Carnegie Mellon University}{bblum@cs.cmu.edu}
\author{Ben Blum}

% tech report stuff
\trnumber{CMU-PDL-16-103}
\keywords{concurrency, stateless model checking, data race analysis, verification}

\abstract{
	\renewcommand{\baselinestretch}{1.0}
	\small
	\em
	%We have developed the
	The Iterative Deepening algorithm allows stateless model checkers to adjust preemption points on-the-fly.
	%incorporate dynamic data-race analysis during testing.
	%The algorithm
	It uses dynamic data-race detection to avoid necessarily preempting on every shared memory access,
	and ignores false-positive data race candidates arising from certain heap allocation patterns.
	%When an Iterative Deepening test reaches completion,
	An Iterative Deepening test that reaches completion soundly verifies all possible thread interleavings of that test.
}

\support{\small{
	{\bf Acknowledgements:}
	Michael J. ``Sully'' Sullivan, Carlo Angiuli, and Nathaniel Filardo double-checked the proof and provided valuable feedback on notation style and the flow of explanation.
Thanks as well to the anonymous OOPSLA reviewer \#4, whose suggestions helped improve the formality of our definitions.
%This work was supported in part by
We thank
the member companies of the PDL Consortium (Broadcom,
Citadel, EMC, Facebook, Google, Hewlett-Packard Labs, Hitachi, Intel,
Microsoft Research, MongoDB, NetApp, Oracle, Samsung America, Seagate,
Tintri, Two Sigma, Uber, Veritas and Western Digital)
and by the U.S. Army Research Office under grant number W911NF0910273.
}}

\maketitle

\section{Introduction}
In our OOPSLA paper \cite{quicksand} we introduced Iterative Deepening,
an algorithm which combines
stateless model checking \cite{verisoft,dpor}
and
data-race analysis \cite{eraser,hybriddatarace}.
It dynamically adds new state spaces to test each time a data-race candidate is found.
We present here two proofs concerning the technique.

First, we prove that given enough time, Iterative Deepening will converge to the same degree of schedule coverage that would be provided by a na\"{i}ve model checker (MC) preempting on every single instruction.
In other words, when Iterative Deepening finishes generating new data-race PPs, and completes all associated state spaces,
it serves as a full verification of {\em all possible} schedules of the given test case.

Second, we prove the soundness of the {\em malloc-recycle} false-positive-elimination tactic discussed in the paper.
Despite powerful reduction techniques, each state space is still exponentially sized,
%so any way to avoid exploring some in advance is obviously beneficial.
so we wish to avoid exploring any we could prove to be redundant.
We prove that it is safe to eliminate malloc-recycle data race candidates immediately, %upon discovering them,
without bothering to explore their associated state spaces.

This tech report is supplemental material to our main paper; we assume the reader is already familiar with our motivation and terminology.

\input{di}
\input{assumptions}
\input{convergence}
\input{recycle}

\bibliographystyle{abbrvnat}
\bibliography{../citations}{}
\end{document}